---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r}
data <- read.csv(unz("activity.zip","activity.csv"),
                 header=TRUE,
                 na.strings="NA",
                 colClasses=c("numeric","character","numeric"))
data$DateTime <- strptime(paste(data$date, sprintf("%04d",data$interval), sep=" "),
                          format="%Y-%m-%d %H%M")
```

## What is mean total number of steps taken per day?
```{r}
stepsPerDay <- aggregate(steps~date,data,FUN=sum)
hist(stepsPerDay$steps, xlab="Steps Per Day", breaks=15, main="Histogram of Steps Per Day")
mean(stepsPerDay$steps)
median(stepsPerDay$steps)
```

## What is the average daily activity pattern?
```{r}
stepsPerFive <- aggregate(steps~interval,data,FUN=mean)
plot(stepsPerFive$interval,stepsPerFive$steps,
     type="l",
     xlab="Interval",
     ylab="Mean number of steps")
```
Find the interval with the maximum number of average steps
```{r}
stepsPerFive[stepsPerFive$steps == max(stepsPerFive$steps),]$interval
```

## Imputing missing values
Calculate total number of missing values
```{r}
sum(is.na(data$steps))
```
We want to replace the NA values with the mean for that interval, which we calculated above
```{r}
noData <- data[is.na(data$steps),]
replacementData <- stepsPerFive[stepsPerFive$interval == noData$interval,]
data$steps <- replace(data$steps,
                      is.na(data$steps), 
                      replacementData$steps)
```
Make a new histogram with the imputed values as well as calculating the new mean and median
```{r}
newStepsPerDay <- aggregate(steps~date,data,FUN=sum)
hist(newStepsPerDay$steps, xlab="Steps Per Day", 
     breaks=15, main="Histogram of Steps Per Day with Imputed Values")
mean(newStepsPerDay$steps)
median(newStepsPerDay$steps)
```
Clearly, there is very little difference whether we impute the missing values or simply ignore them.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
data$DayOfWeek = factor(ifelse(weekdays(data$DateTime)=="Saturday" |
                                 weekdays(data$DateTime)=="Sunday",
                               "weekend",
                               "weekday"))
weekendData <- data[data$DayOfWeek=="weekend",]
weekdayData <- data[data$DayOfWeek=="weekday",]
stepsPerFiveWeekend <- aggregate(steps~interval,weekendData,FUN=mean)
stepsPerFiveWeekend$DayOfWeek <- "weekend"
stepsPerFiveWeekday <- aggregate(steps~interval,weekdayData,FUN=mean)
stepsPerFiveWeekday$DayOfWeek <- "weekday"
stepsPerFiveTotal <- rbind(stepsPerFiveWeekend,stepsPerFiveWeekday)
xyplot(steps~interval | DayOfWeek, 
       data=stepsPerFiveTotal, 
       ylab="Number of steps", 
       xlab="Interval",
       layout=c(1,2), 
       type="l")
```
